---
title: "Access data in the Australian Mental Health Systems Models Dataverse repository."
authors: 
- Matthew Hamilton
date: 2022-08-28T21:13:14-05:00
banner: img/banners/openminddv.png
categories: 
- Resources for Modellers
- Tutorials
tags: 
- R
- Repositories
- Repositories (Data)
output: hugodown::md_document
autonumbering: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# 1. Objectives

In this tutorial we will:

1. Briefly describe the concepts relating to the Australian Mental Health Systems Models Dataverse repository; and

1. Outline the steps required to search for and download files contained in repository datasets using two alternative approaches;
    - Using a web based interface (very simple, suitable for low volume requests where reproducibility is not required); and
    - Using R commands (an approach that requires a little more up-front learning, but which can be used in automated and reproducible workflows)


# 2. Concepts

Information about how to install.

```{r cars}
summary(cars)

fit <- lm(dist ~ speed, data = cars)
```



```{r}
fit
```

# Worked example

Some simple code to demonstrate how to use it.

```{r pie, fig.cap='A fancy pie chart.', tidy=FALSE}
par(mar = c(0, 1, 0, 1))
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```
